---
title: "Explore the association between age, CTCF, and TP53 and other variables"
output:
  html_document:
  df_print: paged
---

# Introduction

Many dynamic eQTLs conditioning on age, or CTCF, and TP53 disappear 
after including a few PEER factors into the model. Here we want to
explore the underlying reasons. 
  
# Load required library
  
```{r warning=FALSE, echo = TRUE, results = 'hide', warning = FALSE, message = FALSE}
library(reshape2)
library(ggplot2)
```

# Read in immune cell composition data

```{r}
imc = read.table("data/CIBERSORTx_Results.txt", header = T, sep="\t", 
                 as.is = T, check.names = F)
dim(imc)
imc[1:2,1:3]
colnames(imc)

rownames(imc) = imc$Mixture
imc = imc[, -c(1, 24:26)]
dim(imc)
imc[1:2,1:3]

refic = which.max(apply(imc, 2, median))
refic

ref = imc[, refic]
log_imc = log((imc[,-refic] + 10^-6)/(ref + 10^-6)) 
log_imc[1:5,]
log_imc = data.frame(log_imc)
```

# Read in gene expression data

```{r fig.height=3, fig.width=3}
edat = read.table("data/Whole_Blood_trec.txt.gz")
dim(edat)
edat[1:2,1:3]
names(edat) = gsub(".", "-", names(edat), fixed=TRUE)
dim(edat)
edat[1:2,1:3]

n10 = rowSums(edat > 10)/ncol(edat)
n20 = rowSums(edat > 20)/ncol(edat)
table(n10 > 0.20)
table(n10 > 0.25)
table(n20 > 0.20)
table(n20 > 0.25)

edat = edat[which(n10 > 0.25),]
dim(edat)

rd  = apply(edat, 2, quantile, prob=0.75)
rd0 = colSums(edat)

par(mar=c(5,4,1,1), bty="n")
plot(rd0, rd, xlab="total read counts per sample", ylab="75th percentile")

table(rownames(imc) == names(edat))

g1 = grep("ENSG00000141510", rownames(edat))
g1
TP53 = unlist(edat[g1,])
summary(TP53)

g1 = grep("ENSG00000102974", rownames(edat))
g1
CTCF = unlist(edat[g1,])
summary(CTCF)

TP53 = (TP53/rd0)*median(rd0)
CTCF = (CTCF/rd0)*median(rd0)

summary(TP53)
summary(CTCF)
```

# Read in PEER factors and PCs

```{r}
cdat = read.table("data/Whole_Blood.v8.covariates.txt", header = T, 
                        sep ='\t', as.is = T, check.names = F)
dim(cdat)
cdat[1:3,1:2]
cdat$ID[1:10]

rownames(cdat) = cdat$ID
cdat = data.matrix(cdat[c(1:2,6:10), -1])
dim(cdat)
cdat[1:3,1:2]

cdat = data.frame(t(cdat))
dim(cdat)
cdat[1:3,1:2]

table(rownames(cdat) == names(edat))
```

# Read in covariate data
```{r}
xdat = read.table("data/GTEx_Analysis_v8_Annotations_SubjectPhenotypesDS.txt", 
                  as.is=TRUE, sep="\t", header=TRUE)
dim(xdat)
xdat[1:2,]

xdat = xdat[match(names(edat), xdat$SUBJID),]
dim(xdat)
xdat[1:2,]
table(rownames(imc) == xdat$SUBJID)

age = strsplit(xdat$AGE, split="-")
age[1:2]
table(sapply(age, length))

age = sapply(age, function(v){mean(as.numeric(v))})
table(age)
xdat$age = round(age+0.01)
table(xdat$age)
```


# Read in the data used in dynamic eQTL analysis to double check
```{r fig.height=4, fig.width=4}
par(mar=c(5,4,1,1), bty="n")

dat_dyn = read.csv("data/Whole_Blood_glm_covariates0.csv")
dim(dat_dyn)
dat_dyn[1:2,]

table(dat_dyn$SUBJID == xdat$SUBJID)
setequal(dat_dyn$SUBJID,xdat$SUBJID)
mat1 = match(xdat$SUBJID, dat_dyn$SUBJID)
dat_dyn = dat_dyn[mat1,]
table(dat_dyn$SUBJID == xdat$SUBJID)

cor(dat_dyn$tp53, log10(TP53))
plot(dat_dyn$tp53, log10(TP53), cex=0.5)

cor(dat_dyn$ctcf, log10(CTCF))
plot(dat_dyn$ctcf, log10(CTCF), cex=0.5)

table(xdat$age,dat_dyn$age)
```

# check associations for age
```{r fig.height=3, fig.width=4}

par(mar=c(5,4,1,1), bty="n")
boxplot(imc$Neutrophils ~ xdat$age, ylab="Neutrophils", xlab="Age")
```

# check associations for CTCF
```{r fig.height=3, fig.width=3}
par(mar=c(5,4,1,1), bty="n")

cr1 = cor(imc, CTCF, method="spearman")
df1 = data.frame(cell_type = rownames(cr1), cr_CTCF=round(cr1,3))
rownames(df1) = NULL
df1 = df1[order(df1$cr_CTCF),]
df1

cr2 = cor(cdat, CTCF, method="spearman")
df2 = data.frame(variables = rownames(cr2), cr_CTCF=round(cr2,3))
rownames(df2) = NULL
df2

plot(imc$`NK cells activated`, CTCF, xlab="NK cells activated", cex=0.5)
plot(imc$`B cells memory`, CTCF, xlab="B cells memory", cex=0.5)

plot(cdat$InferredCov4, log10(CTCF), xlab="PEER factor 4", cex=0.5)
plot(cdat$InferredCov5, log10(CTCF), xlab="PEER factor 5", cex=0.5)

fit0 = lm(CTCF ~ ., data = log_imc)
summary(fit0)

df0 = data.frame(fitted=fitted(fit0), CTCF)
dim(df0)
df0[1:2,]

plot(df0$fitted, df0$CTCF, xlab = 'Fitted values',
     ylab = "CTCF", main = '', pch = 19, cex=0.5)

fit1 = lm(CTCF ~ ., data = cdat)
summary(fit1)

df1 = data.frame(fitted=fitted(fit1), CTCF)
dim(df1)
df1[1:2,]

plot(df1$fitted, df1$CTCF, xlab = 'Fitted values',
     ylab = "CTCF", main = '', pch = 19, cex=0.5)
```

# check associations for TP53

```{r fig.height=3, fig.width=3}
par(mar=c(5,4,1,1), bty="n")

cr1 = cor(imc, TP53, method="spearman")
df1 = data.frame(cell_type = rownames(cr1), cr1)
rownames(df1) = NULL
df1 = df1[order(df1$cr1),]
df1

fit1 = lm(TP53 ~ ., data = log_imc)
summary(fit1)

df1 = data.frame(fitted=fitted(fit1), TP53)
dim(df1)
df1[1:2,]

plot(df1$fitted, df1$TP53, xlab = 'Fitted values',
     ylab = "TP53", main = '', pch = 19, cex=0.5)

plot(imc$`B cells memory`, TP53, xlab="B cells memory", cex=0.5)
plot(imc$Neutrophils, TP53, xlab="Neutrophils", cex=0.5)

cr2 = cor(cdat, TP53, method="spearman")
df2 = data.frame(variables = rownames(cr2), cr_TP53=round(cr2,3))
rownames(df2) = NULL
df2

plot(cdat$InferredCov2, log10(TP53), xlab="PEER factor 2", cex=0.5)
plot(cdat$InferredCov3, log10(TP53), xlab="PEER factor 3", cex=0.5)

fit1 = lm(TP53 ~ ., data = cdat)
summary(fit1)

df1 = data.frame(fitted=fitted(fit1), TP53)
dim(df1)
df1[1:2,]

plot(df1$fitted, df1$TP53, xlab = 'Fitted values',
     ylab = "TP53", main = '', pch = 19, cex=0.5)
```

```{r}
gc()
sessionInfo()
```
