---
title: "Linear model fit for p-value discrepancies"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=TRUE}
rasq = read.csv("linear_model_cov.csv")
rasq = rasq[!is.na(rasq$pvalT),]
dim(rasq)
rasq[1:2,]

rasq$qvalR = 10^rasq$qval
rasq$qvalT[rasq$qvalT<1e-100] = 1e-100
rasq$qvalR[rasq$qvalR<1e-100] = 1e-100
rasq$pvalT[rasq$pvalT<1e-100] = 1e-100
rasq$pvalR[rasq$pvalR<1e-100] = 1e-100
dim(rasq)

#rasq$rnlt = -log10(rasq$qvalR)
#rasq$tnlt = -log10(rasq$qvalT)
rasq$rnlt = -log10(rasq$pvalR)
rasq$tnlt = -log10(rasq$pvalT)

cent = function(x){
  med = median(x)
  (x-med)
}
norm = function(x){
  sd = sd(x)
  x/sd
}

odNB = rasq$odNB; odNB[odNB<1e-3]=1e-3; odNB[odNB>8]=8; 
odBB = rasq$odBB; odBB[odBB<1e-3]=1e-3; odBB[odBB>8]=8
odR = rasq$OD
lodNBp = log10(odNB); 
lodBBp = log10(odBB); 
lodRp = -log10(odR);
lodNB = cent(norm(lodNBp)); 
lodBB = cent(norm(lodBBp)); 
lodR = cent(norm(lodRp));



hiR = function(cuts, y){
   kp = abs(y)>cuts
   c(mean(y[kp]>0), sum(kp)) 
}
#show RASQUAL 
GfSNP = rasq$NfSNP
GfSNP[rasq$NfSNP>1]=3
GfSNP[rasq$NfSNP>2]=6
GfSNP[rasq$NfSNP>4]=12
GfSNP[rasq$NfSNP>8]=24
GfSNP[rasq$NfSNP>16]=48
cutoff = 25
rasq$rnlt[rasq$rnlt>cutoff] = cutoff
rasq$tnlt[rasq$tnlt>cutoff] = cutoff
y = rasq$rnlt-rasq$tnlt
z = y>0
kp0 = abs(y)>=0
kp1 = abs(y)>=1
kp5 = abs(y)>=5
kp10 = abs(y)>=10
kp15 = abs(y)>=15

ag0 = aggregate(z[kp0], by=list(GfSNP[kp0]), FUN=mean)
ag1 = aggregate(z[kp1], by=list(GfSNP[kp1]), FUN=mean)
ag5 = aggregate(z[kp5], by=list(GfSNP[kp5]), FUN=mean)
ag10 = aggregate(z[kp10], by=list(GfSNP[kp10]), FUN=mean)
ag15 = aggregate(z[kp15], by=list(GfSNP[kp15]), FUN=mean)

cg0 = aggregate(rep(1,sum(kp0)), by=list(GfSNP[kp0]), FUN=sum)
cg1 = aggregate(rep(1,sum(kp1)), by=list(GfSNP[kp1]), FUN=sum)
cg5 = aggregate(rep(1,sum(kp5)), by=list(GfSNP[kp5]), FUN=sum)
cg10 = aggregate(rep(1,sum(kp10)), by=list(GfSNP[kp10]), FUN=sum)
cg15 = aggregate(rep(1,sum(kp15)), by=list(GfSNP[kp15]), FUN=sum)
ind = 1:15
frR = sapply(ind, hiR, y=y)

numg = cors = cors2 = odnb = rep(0, length(ind))
for(i in ind){
  kpi = abs(y)>i
  numg[i] = sum(kpi)
  cors[i] = round((cor(lodR[kpi], lodNB[kpi])),2)
  cors2[i] = round((cor(lodR[kpi], lodBB[kpi])),2)
  odnb[i] = mean(lodNB[kpi])
}

```

## R Markdown

Plotting fraction of RASQUAL more significant than TReCASE and over-dispersion correlations vs dicrepancy in p-value and number of F SNPs

```{r fig1, fig.height=8, fig.width=8, fig.align="center", echo=TRUE }
cexes = 1.8
par(mfrow=c(2,2))
par(mar=c(5,5,3,1))

plot(ind, frR[1,], xlab="|log10(T p-val)-log10(R p-val)|", ylab="%p(R)<p(T)", 
main="", bty="n", cex.main=cexes, cex.lab=cexes, cex.axis=cexes, cex=log10(frR[2,]))
legend("bottomright", legend=c("#genes", 10^(4:2)), pt.cex=c(1, 4:2), 
       cex=cexes, bty="n", pch=c(NA, 1,1,1))
legend("topleft", legend=c(2500, 500, 100), pt.cex=log10(c(2500, 500, 100)), 
       cex=cexes, bty="n", pch=c(1,1,1))
axis(1, at=c(12), 12, cex.axis=cexes)

plot(ind, odnb, xlab="|log10(T p-val)-log10(R p-val)|", ylab="normalized log(NBod)", 
main="", bty="n", cex.main=cexes, cex.lab=cexes, cex.axis=cexes, cex=log10(frR[2,]))
legend("bottomleft", legend=c(2500, 500, 100), pt.cex=log10(c(2500, 500, 100)), 
       cex=cexes, bty="n", pch=c(1,1,1))
axis(1, at=c(12), 12, cex.axis=cexes)


x = 1:6
plot(x, ag0[,2], ylim=c(0,1), type="b", xlab="#fSNP", ylab="%p(R)<p(T)", 
      main="", bty="n", xaxt="n", cex=log10(cg0[,2]), 
      xlim=c(1,6.1), cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
points(x, ag5[,2], col="blue", type="b", pch=2, cex=log10(cg5[,2]))
points(x, ag10[,2], col="goldenrod", type="b", pch=3, cex=log10(cg10[,2]))
points(x, ag15[,2], col="red", type="b", pch=4, cex=log10(cg15[,2]))
axis(1, at=x, ag0[,1], cex.lab=cexes, cex.axis=cexes)
legend("topleft", legend=c("all",">5", ">10", ">15"), 
text.col=c("black", "blue", "goldenrod", "red"), pch=c(1:4), bty="n", cex=cexes)


plot(ind, cors, cex=log10(numg), pch=1, bty="n", xlab="|log10(T p-val)-log10(R p-val)|", 
ylab="OD correlation", main="", cex.main=cexes, cex.lab=cexes, cex.axis=cexes, ylim=c(0,1))
points(ind, cors2, cex=log10(numg), pch=2)
axis(1, at=c(12), 12, cex.axis=cexes)
legend("bottomleft", c("OD(NB)", "OD(BB)"), bty="n", pch=c(1,2), cex=cexes)
legend("bottomright", legend=c("#genes", 10^(4:2)), 
       pt.cex=c(1, 4:2), cex=cexes, bty="n", pch=c(NA, 1,1,1))

```


```{r echo=FALSE}
#prepare covariates

NfSNP = rasq$NfSNP;
NfSNP[rasq$NfSNP>=60] = 60;
NfSNP = log(NfSNP)
NfSNP = cent(norm(NfSNP))

NrSNP = rasq$NrSNP;
NrSNP = log(NrSNP)
NrSNP = cent(norm(NrSNP))
NrSNP[NrSNP< -5] = -5
NrSNP[NrSNP> +5] =  5
#hist(NrSNP)

af = rasq$AF
af = norm(af)

hwec2 = rasq$HWEChisq; hwec2[hwec2<1e-2]=1e-2; hwec2[hwec2>30]=30;hwec2=log(hwec2)
hwec2 = norm(hwec2)

derr = rasq$DeltaErr; derr[derr<1e-10]=1e-10; derr[derr>0.01]=0.01
derr = norm(derr)
phib = abs(rasq$PhiBias-.5); phib[phib>.1]=.1
phib = norm(phib)

asT = rasq$asT
asT = cent(norm(log(asT)))

asR = cent(norm(log(rasq$asR)))

cutoff = 15
rasq$rnlt[rasq$rnlt>cutoff] = cutoff
rasq$tnlt[rasq$tnlt>cutoff] = cutoff
y = rasq$rnlt-rasq$tnlt


mpp = rasq$mpp
mpp[mpp<1e-5] = 1e-5
mpp = norm(mpp)

#interactions
i12 = asR*NfSNP
i13 = asR*lodBB
i23 = NfSNP*lodBB
i123 = asR*lodBB*NfSNP
#i123[i123< -10] = -10
#i123[i123>  10] = 10
ia14 = asT*asR
```

Plotting Over-dispersions

```{r fig2, fig.height=8, fig.width=12, fig.align="center", echo=TRUE }
par(mfrow=c(2,3))
par(mar=c(5,5,3,1))

plot(density(lodNB), bty="n", main="log10(NB OD)", 
     xlab="log(od)", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
plot(density(lodBB), bty="n", main="log10(BB OD)", 
     xlab="log(od)", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
plot(density(lodR), bty="n", main="log10(RASQ OD)",
     xlab="log(od)", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
plot(lodNB, lodBB, main="NB vs BB over-disp", pch=20, cex=0.5, col=rgb(0.8,0.1,0.1,0.5), bty="n",
     xlab="log(NB OD)", ylab="log(BB OD)", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
abline(a=0, b=1)
plot(lodNB, lodR, main="NB vs Rasq over-disp", pch=20, cex=0.5, col=rgb(0.8,0.1,0.1,0.5), bty="n",
     xlab="log(NB OD)", ylab="log(RASQ OD)", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
abline(a=0, b=1)
plot(lodR, lodBB, main="Rasq vs BB over-disp", pch=20, cex=0.5, col=rgb(0.8,0.1,0.1,0.5), bty="n",
     xlab="log(RASQ OD)", ylab="log(BB OD)", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
abline(a=0, b=1)

```


Plotting RASQUAL read counts, normalized NfSNP and their interaction along with interactions with BB over-dispersion

```{r fig3, fig.height=8, fig.width=12, fig.align="center", echo=TRUE }
par(mfrow=c(2,3))
par(mar=c(5,5,3,1))

plot(density(asR), bty="n", main="#as-RASQ", xlab="#as-RASQUAL", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
plot(density(NfSNP), bty="n", main="#fSNPs", xlab="#fSNPs", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
#plot(density(NrSNP), bty="n", main="#rSNPs", xlab="#fSNPs", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
plot(density(i12), bty="n", main="interac, #asR & #fSNP", 
     xlab="#as-RASQ * #fSNPs", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
plot(density(i13), bty="n", main="interac, #asR & #log(BB OD)", 
     xlab="#asR * log10(BB OD)", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
plot(density(i23), bty="n", main="interac, #fSNP & log(BB OD)", 
     xlab="#fSNPs * log10(BB OD)", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
plot(density(i123), bty="n", main="interac, #asR & #fSNP & log(BB OD0", xlim=c(-10,10),
     xlab="#as-RASQ * #fSNPs * log10(BB OD)", , cex.main=cexes, cex.lab=cexes, cex.axis=cexes)

```

```{r fig4, fig.height=8, fig.width=12, fig.align="center", echo=TRUE }
par(mfrow=c(2,3))
par(mar=c(5,5,3,1))

plot(density(asT), bty="n", main="#as-TReCASE", 
     xlab="#as-TReCASE", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
plot(density(ia14), bty="n", main="interact, #as-RASQ & #as-TReCASE", 
     xlab="#as-RASQ*#as-TReCASE", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
plot(asT, asR, bty="n", main="#as-RASQ vs #as-TReCASE", pch=20, cex=0.5, col=rgb(0.8,0.1,0.1,0.5), bty="n", 
              xlab="#as-TReCASE", ylab="#as-RASQUAL", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)

plot(density(NrSNP), bty="n", main="#rSNPs", 
     xlab="#fSNPs", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
plot(density(y), bty="n", main="difference in log10 p-values", 
     xlab="log10(T p-val)-log10(R p-val)", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
plot(rasq$tnlt, rasq$rnlt, bty="n", main="RASQUAL vs TReCASE p-val", pch=20, cex=0.5, col=rgb(0.8,0.1,0.1,0.5), bty="n", 
              xlab="-log10(T p-val)", ylab="-log10(R p-val)", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
#rasq$rnlt-rasq$tnlt") - negative implies TReCASE is more significant

```

Other covariates: median permuted p-value, mapping error, reference bias and Chi2 of Hardy Weinberg Equilibrium
```{r fig5, fig.height=8, fig.width=8, fig.align="center", echo=TRUE }
par(mfrow=c(2,2))
par(mar=c(5,5,3,1))
plot(density(mpp), bty="n", main="#median permuted p-val", 
     xlab="normalized p-value", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
plot(density(derr), bty="n", main="mapping error", 
     xlab="normalized error", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
plot(density(phib), bty="n", main="#reference bias", 
     xlab="normalized deviation from 0.5", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
plot(density(hwec2), bty="n", main="#HWE equilibrium", 
     xlab="log(Chi2)", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
```


Plotting RASQUAL read counts, normalized NfSNP and their interaction along with interactions with BB over-dispersion

```{r fig6, fig.height=8, fig.width=12, fig.align="center", echo=TRUE }
par(mfrow=c(2,3))
par(mar=c(5,5,3,1))

plot(lodNB, y, main="discrep. vs log(NB)", ylab="diff", pch=20, cex=0.5, col=rgb(0.8,0.1,0.1,0.5), 
     xlab="log(NB OD)", cex.main=cexes,  cex.lab=cexes, cex.axis=cexes, bty="n")
plot(lodBB, y, main="discrep. vs log(BB)", ylab="diff", pch=20, cex=0.5, col=rgb(0.8,0.1,0.1,0.5),  
     xlab="log(BB OD)", cex.main=cexes, cex.lab=cexes, cex.axis=cexes, bty="n")
plot(NfSNP, y, main="discrep. vs #fSNP", ylab="diff", pch=20, cex=0.5, col=rgb(0.8,0.1,0.1,0.5), 
     xlab="num-fSNP", cex.main=cexes, cex.lab=cexes, cex.axis=cexes, bty="n")
plot(asR, y, main="discrep. vs ASReC-RL cnt", ylab="diff",  pch=20, cex=0.5, col=rgb(0.8,0.1,0.1,0.5),
     xlab="ASReC-RL", cex.main=cexes, cex.lab=cexes,cex.axis=cexes, bty="n")
plot(asT, y, main="discrep. vs ASReC-TR cnt", ylab="diff",  pch=20, cex=0.5, col=rgb(0.8,0.1,0.1,0.5),
     xlab="ASReC-TR", cex.main=cexes, cex.lab=cexes, cex.axis=cexes, bty="n")
plot(mpp, y, main="discrep. vs med. per. p-val", ylab="diff",  pch=20, cex=0.5, col=rgb(0.8,0.1,0.1,0.5), 
     xlab="med. per. p-val", cex.main=cexes, cex.lab=cexes, cex.axis=cexes, bty="n")

```


Adding a wrapper to linear model and anova processing to be run with a subset of the genes.
```{r echo=FALSE}
#a function to fit model on a subset of genes

subset_fit = function(kp, digits=3){
  nms = c("log(OD_BB)", "ASReC_R", "n-fSNP", "ASReC_R:n-fSNP", "ASReC_R:log(OD_BB)",
   "n-fSNP:log(OD_BB)", "ASReC_R:n-fSNP:log(OD_BB)", "n-rSNP", "log(OD_NB)",
   "ASReC_T", "ASReC_T:ASReC_R", "AF", "Chi2-HWEC", "Map.Err", "Ref.Bias", "Med(perm-p)") 
  lm.2 = lm(y[kp]~lodBB[kp]+asR[kp]+NfSNP[kp]
             +i12[kp]+i13[kp]+i23[kp]+i123[kp]
             +NrSNP[kp]+lodNB[kp]+asT[kp]+ia14[kp]
             +af[kp]+hwec2[kp]+derr[kp]+phib[kp]+mpp[kp])
  names(lm.2$coefficients)[-1] = nms
  
  
  
  library(car)
  an.1 = anova(lm.2)
  tss = sum(an.1[,2])
  an.3 = Anova(lm.2, type="3")
  ano3 = cbind(round(100*an.3[,1]/tss,digits-2), format(an.3[,4], scientific=T, digits=digits-1))[-1,]
  
  
  lm.m0 = lm(y[kp]~1)
  lm.m1 = lm(y[kp]~lodBB[kp])
  lm.m2 = lm(y[kp]~asR[kp])
  lm.m3 = lm(y[kp]~NfSNP[kp])
  lm.m4 = lm(y[kp]~i12[kp])
  lm.m5 = lm(y[kp]~i13[kp])
  lm.m6 = lm(y[kp]~i23[kp])
  lm.m7 = lm(y[kp]~i123[kp])
  lm.m8 = lm(y[kp]~NrSNP[kp])
  lm.m9 = lm(y[kp]~lodNB[kp])
  lm.m10 = lm(y[kp]~asT[kp])
  lm.m11 = lm(y[kp]~ia14[kp])
  
  lm.m12 = lm(y[kp]~af[kp])
  lm.m13 = lm(y[kp]~hwec2[kp])
  lm.m14 = lm(y[kp]~derr[kp])
  lm.m15 = lm(y[kp]~phib[kp])
  lm.m16 = lm(y[kp]~mpp[kp])
  
  mr2 = 
  c(round(summary(lm.m0)$r.squared*100,digits-2), lm.m0$coef[1], summary(lm.m0)$coefficients[1,4],
  round(summary(lm.m1)$r.squared*100,digits-2), lm.m1$coef[2], summary(lm.m1)$coefficients[2,4],
  round(summary(lm.m2)$r.squared*100,digits-2), lm.m2$coef[2], summary(lm.m2)$coefficients[2,4],
  round(summary(lm.m3)$r.squared*100,digits-2), lm.m3$coef[2], summary(lm.m3)$coefficients[2,4],
  round(summary(lm.m4)$r.squared*100,digits-2), lm.m4$coef[2], summary(lm.m4)$coefficients[2,4],
  round(summary(lm.m5)$r.squared*100,digits-2), lm.m5$coef[2], summary(lm.m5)$coefficients[2,4],
  round(summary(lm.m6)$r.squared*100,digits-2), lm.m6$coef[2], summary(lm.m6)$coefficients[2,4],
  round(summary(lm.m7)$r.squared*100,digits-2), lm.m7$coef[2], summary(lm.m7)$coefficients[2,4],
  round(summary(lm.m8)$r.squared*100,digits-2), lm.m8$coef[2], summary(lm.m8)$coefficients[2,4],
  round(summary(lm.m9)$r.squared*100,digits-2), lm.m9$coef[2], summary(lm.m9)$coefficients[2,4],
  round(summary(lm.m10)$r.squared*100,digits-2), lm.m10$coef[2], summary(lm.m10)$coefficients[2,4],
  round(summary(lm.m11)$r.squared*100,digits-2), lm.m11$coef[2], summary(lm.m11)$coefficients[2,4],
  round(summary(lm.m12)$r.squared*100,digits-2), lm.m12$coef[2], summary(lm.m12)$coefficients[2,4],
  round(summary(lm.m13)$r.squared*100,digits-2), lm.m13$coef[2], summary(lm.m13)$coefficients[2,4],
  round(summary(lm.m14)$r.squared*100,digits-2), lm.m14$coef[2], summary(lm.m14)$coefficients[2,4],
  round(summary(lm.m15)$r.squared*100,digits-2), lm.m15$coef[2], summary(lm.m15)$coefficients[2,4],
  round(summary(lm.m16)$r.squared*100,digits-2), lm.m16$coef[2], summary(lm.m16)$coefficients[2,4])
  mr2 = t(matrix(mr2, nrow=3))
  
  tab = summary(lm.2)$coef
  tab[,1] = round(tab[,1],digits)
  tab[,2] = round(tab[,2],digits)
  tab = tab[,-3]
  tab = cbind(tab, mr2[,2], mr2[,3])
  tab[,3] = format(as.numeric(tab[,3]), scientific=T, digits=2)
  tab[,4] = round(as.numeric(tab[,4]), digits)
  tab[,5] = format(as.numeric(tab[,5]), scientific=T, digits=2)
  colnames(tab) = c("Est", "SE", "P-val", "Marg.Est", "Marg.P")
  #tab
  
  ano = an.1
  ano[,3] = round(100*ano[,2]/tss,digits-2)
  ano[,5] = format(ano[,5], scientific=T, digits=2)
  ano[,2] = round(ano[,2])
  ano = cbind(ano[,-c(1,4)], ano3)
  ano = ano[-17,]
  ano = cbind(ano, mr2[-1,1])
  rownames(ano) = names(lm.2$coefficients)[-1]
  colnames(ano)=c("SumSq", "Typ1per", "T1P-val", "Typ3per", "T3P-val", "Marg.R")
  #ano
  list(tab=tab, ano=ano, r2=summary(lm.2)$r.squared)
}


```

Fitting all 13978 genes



```{r echo=TRUE}
#with several cutoffs
#kp = abs(y)>=(cutoff-5)
cutsub = 0
kp = abs(y)>=cutsub
sum(kp)
resA = subset_fit(kp, digits=3)

resA$tab
resA$ano
resA$r2

write.csv(resA$ano, sprintf("anova_cut%s.csv", cutsub), quote=F)
write.csv(resA$tab, sprintf("model_cut%s.csv", cutsub), quote=F)

```

Less stringent cutoff (1202 genes)
```{r echo=TRUE}
cutsub = 5
kp = abs(y)>=cutsub
sum(kp)
resA = subset_fit(kp)

resA$tab
resA$ano
resA$r2

write.csv(resA$ano, sprintf("anova_cut%s.csv", cutsub), quote=F)
write.csv(resA$tab, sprintf("model_cut%s.csv", cutsub), quote=F)

```
The most discrepant 221 genes

```{r echo=TRUE}
cutsub = cutoff-5
kp = abs(y)>=cutsub
sum(kp)
resA = subset_fit(kp)

resA$tab
resA$ano
resA$r2

write.csv(resA$ano, sprintf("anova_cut%s.csv", cutsub), quote=F)
write.csv(resA$tab, sprintf("model_cut%s.csv", cutsub), quote=F)

```
plot(lodNB[kp], y[kp], main="discrep. vs log(NB)", ylab="diff", cex=0.5, 
     col=rgb(0.8,0.1,0.1,0.5), xlab="log(NB OD)", cex.main=cexes, 
     cex.lab=cexes, cex.axis=cexes, pch=20)


Keeping only genes with difference <10
```{r echo=TRUE}
cutsub = 10
kp = abs(y)<cutsub
sum(kp)
resA = subset_fit(kp)

resA$tab
resA$ano
resA$r2

write.csv(resA$ano, sprintf("anova_less%s.csv", cutsub), quote=F)
write.csv(resA$tab, sprintf("model_less%s.csv", cutsub), quote=F)
```

Keeping only genes with difference <5
```{r echo=TRUE}
cutsub = 5
kp = abs(y)<cutsub
sum(kp)
resA = subset_fit(kp)

resA$tab
resA$ano
resA$r2

write.csv(resA$ano, sprintf("anova_less%s.csv", cutsub), quote=F)
write.csv(resA$tab, sprintf("model_less%s.csv", cutsub), quote=F)

```


```{r od, echo=TRUE}
kp = abs(y)>5
lm.r = lm(lodR[kp]~lodBB[kp]+lodNB[kp]+lodBB[kp]:lodNB[kp])
odr = lm.r$residuals
lm.3 = lm(y[kp]~lodBB[kp]+asR[kp]+NfSNP[kp]
             +i12[kp]+i13[kp]+i23[kp]+i123[kp]
             +NrSNP[kp]+lodNB[kp]+asT[kp]+ia14[kp]+odr
             +af[kp]+hwec2[kp]+derr[kp]+phib[kp]+mpp[kp])
summary(lm.3)
anova(lm.r)
anova(lm.3)

```
```{r fig_last, echo=TRUE}
cexes = 1.2
png("new_od_plot.png", height=4, width=4, res=300, units="in")
par(mar=c(5,5,3,1))
plot(lodNBp, lodBBp, main="NB vs BB over-disp", pch=20, cex=.5, col=rgb(0.4,0.4,0.4,0.5), bty="n",
     xlab="log10(NB overdispersion)", ylab="log(BB overdispersion)", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
abline(a=0, b=1, col="red", lwd=2, lty=3)
dev.off()

rasqS = trecS = rep(0, nrow(rasq))
rasqS[rasq$qvalR<5e-2] = 1
rasqS[rasq$qvalR<1e-3] = 2
rasqS[rasq$qvalR<1e-4] = 3
rasqS[rasq$qvalR<1e-5] = 4

trecS[rasq$qvalT<5e-2] = 1
trecS[rasq$qvalT<1e-3] = 2
trecS[rasq$qvalT<1e-4] = 3
trecS[rasq$qvalT<1e-5] = 4

table(rasqS, trecS)
```


```{r fig_last, echo=TRUE}
cexes = 1.7
png("od_vs_rasq_plot.png", height=4, width=8, res=300, units="in")
par(mfrow=c(1,2))
par(mar=c(5,5,3,1))
plot(lodRp, lodNBp, main="(a) RASQUAL vs NB", pch=20, cex=.5, col=rgb(0.4,0.4,0.4,0.5), bty="n",
     xlab="log10(RASQUAL od)", ylab="log10(TReCASE NB od)", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
abline(a=0, b=1, col="red", lwd=2, lty=3)

plot(lodRp, lodBBp, main="(b) RASQUAL vs BB", pch=20, cex=.5, col=rgb(0.4,0.4,0.4,0.5), bty="n",
     xlab="log10(RASQUAL od)", ylab="log10(TReCASE BB od)", cex.main=cexes, cex.lab=cexes, cex.axis=cexes)
abline(a=0, b=1, col="red", lwd=2, lty=3)
dev.off()
```
